$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-07-23 paper/paper.tex
More involved example

% Constructors
\newcommand{\PO}{\constr{\mathbb{0}}}
\newcommand{\comp}[2]{#1 \, \constr{\parallel} \, #2}
\newcommand{\new}{\constr{\boldsymbol{\nu}} \,}
\newcommand{\send}[2]{#1 \, \constr{\langle} \, #2 \,\constr{\rangle} \,}
\newcommand{\recv}[2]{#1 \, \constr{\mathbb{(}} \, #2 \, \constr{\mathbb{)}} \,}
\newcommand{\suc}{\constr{\scriptstyle 1+}}
\newcommand{\unit}{\constr{\mathbbm{1}}}
\newcommand{\base}[1]{\constr{B[} \, #1 \, \constr{]}}
\newcommand{\channel}[2]{\constr{C[} \, #1 \, \constr{,} \, #2 \, \constr{]}}
\newcommand{\comma}{\, \constr{,} \,}

@@

\begin{example}[Typing derivation]
  \label{example-derivations}
  In \autoref{lst:derivation} we illustrate a typing derivation of the process $\func{p}$ (lines 1-4) using linear and shared types.

  Let $\func{\omega}$ be a shortcut for $\one \comma \one$ in an algebra of shared types.
  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $\zero \comma \zero$.
  To type $\func{p}$, we need a channel over which we can transmit another channel.
  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
  The left-hand process (line 3) is typed following its syntax (line 8).
  For the right-hand (line 4) we need the channel that is going to be sent (line 9).
  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
  Its usage annotation is $\one \comma \one$: we will transmit $\one \comma \zero$ and save $\zero \comma \one$ for the continuation process.

  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing a process that sends part of one channel over another.]
  $\func{p}$ : $\Process_{\suc \constr{zero}}$
  $\func{p}$ = $\new{}$
       $(\comp{\recv{\constr{0}}{} (\recv{\constr{0}}{} \PO)}$
       ${\new{} (\send{\suc \; \constr{0}} {\constr{0}} (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO)})$

  $\func{\_}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\func{p}}{\func{\epsilon}}$
  $\func{\_}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; \one$
       $(\comp{\recv{\constr{0}}{} \; (\recv{\constr{0}}{} \; \PO)}$
       ${\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; (\send{\constr{0}}{\suc \; \suc \; \constr{0}} \; \PO))}$
  \end{lstlisting}
\end{example}

>>>>

% Constructors
\newcommand{\PO}{\constr{\mathbb{0}}}
\newcommand{\comp}[2]{#1 \, \constr{\parallel} \, #2}
\newcommand{\new}{\constr{\boldsymbol{\nu}} \,}
\newcommand{\send}[2]{#1 \, \constr{\langle} \, #2 \,\constr{\rangle} \,}
\newcommand{\recv}[2]{#1 \, \constr{\mathbb{(}} \, #2 \, \constr{\mathbb{)}} \,}
\newcommand{\suc}{\constr{\scriptstyle 1+}}
\newcommand{\unit}{\constr{\mathbbm{1}}}
\newcommand{\base}[1]{\constr{B[} \, #1 \, \constr{]}}
\newcommand{\channel}[2]{\constr{C[} \, #1 \, \constr{;} \, #2 \, \constr{]}}
\newcommand{\comma}{\, \constr{,} \,}

@@

\begin{example}[Typing derivation]
  \label{example-derivations}
  In \autoref{lst:derivation} we illustrate a typing derivation using shared, graded and linear types.
  We use $\func{\omega}$ as a shortcut for $\one \comma \one$ in an algebra of shared types.

  \begin{lstlisting}[label=lst:derivation,mathescape,numbers=left,caption=Typing two processes that send parts of one channel over another.]
$\func{P}$ : $\Process_{\suc \suc \constr{0}}$
$\func{P}$ = $\comp{(\new{} \; (\send{\suc \; \constr{0}} {\constr{0}} \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))}{\recv{\constr{0}}{\recv{\constr{0}}{}} \PO}$

$\func{\gamma}$ : $\PreCtx_{\suc \suc \constr{0}}$
$\func{\gamma}$ = $\constr{[]} \comma \unit \comma \channel{\channel{\unit}{\func{\omega}}}{\li}$

$\func{\vdash P}$ : $\forall \{n\} \to \types{\func{\gamma}}{\constr{[]} \comma \func{\omega} \comma (\suc n \comma \suc n)}{\func{P}}{\constr{[]} \comma \func{\omega} \comma (n \comma n)}$
$\func{\vdash P}$ = $\comp{(\new \; \unit \; \func{\omega} \; \one \; (\send{\suc \; \constr{0}}{\constr{0}} \; \send{\constr{0}}{\suc \; \suc \; \constr{0}} \PO))} {\recv{\constr{0}}{} \; \recv{\constr{0}}{} \PO}$

$\func{\vdash PP}$ : $\types{\constr{[]} \comma \unit}{\constr{[]} \comma \func{\omega}}{\new{} \comp{\func{P}}{\func{P}}}{\constr{[]} \comma \func{\omega}}$
$\func{\vdash PP}$ = $\new \; \channel{\unit}{\func{\omega}} \; \li \; (\suc \suc \constr{0}) \; (\comp{\vdash P}{\vdash P})$
  \end{lstlisting}

  Notice that the \emph{only} difference between the definition of the untyped process $\func{P}$ and its typing derivation are the extra type annotations that accompany scope restriction.
  This extra information requires a (limited) choice.
  To type $\func{P}$, we need a channel over which we can transmit another channel.
  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 9).
  The left-hand process (line 3) is typed following its syntax (line 10).
  For the right-hand (lines 4-5) we need the channel that is going to be sent (lines 11-12).
  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
  Its usage annotation is $\one \comma \one$: we will transmit $\one \comma \zero$ and save $\zero \comma \one$ for the continuation process.

\end{example}
